---
title: "Git and GitHub Tutorial"
author: "ZMT"
date: "01/09/2021"
output: 
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a hands-on tutorial on Git and GitHub. It covers the following topics:

-   Set up
-   Basic bash commands
-   Git commands
-   GitHub operations
-   Data analysis workflow using Git and GitHub

# Set up

We will use the following software in this tutorial.

### Git

Git is the most popular version control system. It allows us to track the changes of texts as a project develops.

Git is pre-installed in Mac OS. Make sure your git version is 2.23.0 or above. You can check your git version in your terminal. First, press Command + Spacebar to open a search bar on your screen. Then, type "terminal" and double click "Terminal" in the left sidebar to open the Terminal. In it, type `git --version` to find out your git version. You can download and update git from [here](https://sourceforge.net/projects/git-osx-installer/).

In Windows, download and install Git from [here](https://git-scm.com/download/win). After the installation, press Windows key and type "git bash". You should see Git Bash in the list. Click on it to open the Git Bash terminal.

### Visual Studio Code

We will edit texts using Visual Studio Code editor. Download and install it [here](https://code.visualstudio.com/).

### GitKraken

GitKraken is a popular Git graphic user interface (GUI) which allow people to use Git without using the command line.

**Pros of GUI**

-   Very low barrier of entry with much better user experience compared to command line
-   Some prefer GUI, including those who can use the command line

**Cons of GUI**

-   The inner-workings of Git are hidden away with GUIs.
-   It leads to dependence on a particular piece of software.
-   Interfaces are different between various GUIs.

**Pros of the command line**

-   Git is a command line tool. All the documentation and resources online refer to the command line.
-   It becomes way more efficient once you grow comfortable with it.
-   Advanced Git features are only available in the command line.
-   The commands are the same across platforms (Windows, Mac OS, Linux).

**Cons of the command line**

-   The learning curve can be steep for beginners.  

Other popular GUIs include:

-   Github Desktop
-   SourceTree
-   Tower

[Here](https://git-scm.com/downloads/guis) is a list of Git GUI clients.

### GitHub account

If you have not done so, sign up for a [GitHub](https://github.com/) account.

# Bash commands

This tutorial mainly focuses on the command line. Before learning Git, we will learn how to navigate between directories and manage files in the Bash command line. It is not the most exciting thing to learn but we will just quick go over some of the commonly used commands.

### Navigation

Open current directory in Explorer (in Windows) or in Finder (in Mac).

```{r eval=FALSE}
# Windows
start . 

# Mac
open .
```

Check the current directory.

```{r eval=FALSE}
pwd
```

List files and directories in the current directory.

```{r eval=FALSE}
ls
ls -a
```

Go to a sub-directory.

```{r eval=FALSE}
cd sub-directory
```

Go to the parent directory.

```{r eval=FALSE}
cd ..
```

### Working with files and directories

Create a file.

```{r eval=FALSE}
touch file.txt
```

Create multiple files.

```{r eval=FALSE}
touch file1.txt file2.txt file3.txt
```

Create a directory.

```{r eval=FALSE}
mkdir directory_name
```

Create multiple directories.

```{r eval=FALSE}
mkdir dir1 dir2 dir3
```

Delete a file.

```{r eval=FALSE}
rm file.txt
```

Delete a directory.

```{r eval=FALSE}
rm -rf directory_name
rmdir directory_name
```

# Git commands

### Configuration

It is essential to add username and email for logging purposes.

```{r eval=FALSE}
git config --global user.name "your_name"
git config --global user.email "your_email"
```

Set Visual Studio Code as the default editor.

```{r eval=FALSE}
git config --global core.editor "code --wait"
```

### Git basics

[pic: initiate, stage, commit]

Start a git repository.

```{r eval=FALSE}
git init
```

Check status.

```{r eval=FALSE}
git status
```

Stage files and directories.

```{r eval=FALSE}
# Stage one or more selected files
git add file.txt file.R

# Stage all files
git add .
```

### Committing


### Branching


### Merging


### Diffs


### Undoing stuffs


# Github operations

### Basics

### Fetch and pull


# Data analysis workflow

### Centralised workflow


### Feature branches


### Pull requests


